\section{SODA-QFT: Semantic Dynamics and Multi-Field Quantum Evolution}
\label{sec:soda_qft}

The final extension of the SODA framework integrates Quantum Field Theory (QFT) and Semantic Information Theory, resulting in the \textbf{SODA-QFT engine}: a multi-field, semantically grounded, GPU-accelerated evolutionary system. Each Nucleus no longer carries a single complex state $\psi_j$, but instead a \textit{spinor-valued quantum field}:

\[
\Psi_j = 
\begin{bmatrix}
\psi^{(1)}_j \\
\psi^{(2)}_j \\
\vdots \\
\psi^{(F)}_j
\end{bmatrix}
\in \mathbb{C}^F,
\]

where each component corresponds to a structural modality (Chaos, Bayesian, Spiking-Neural, etc.), enabling multi-perspective reasoning and cross-field information transfer.

\subsection{Multi-Field Quantum Kuramoto Map}

The Quantum Kuramoto Map (QKM) is generalized to account for cross-field interactions through the multi-field coupling tensor $J_{jk}^{(a,b)}$:

\[
\Psi_j(t+1) 
= 
\frac{1}{Z_j}
\left[
\Psi_j(t) 
+ \sum_{k=1}^{N}
\sum_{b=1}^{F}
J_{jk}^{(a,b)} 
\, e^{i(\phi_k^{(b)} - \phi_j^{(a)})}
\, \psi_k^{(b)}(t)
+ \Delta \omega_j \Psi_j(t)
\right].
\]

This coupling tensor encodes three distinct effects:

\begin{enumerate}
    \item \textbf{Fitness Proximity}: $e^{-|f_j - f_k| / T}$ ensures poorly aligned solutions are de-emphasized.
    \item \textbf{Entropy Polarity}: $e^{-|H_j - H_k|}$ encourages exploration when diversity is high.
    \item \textbf{Semantic Interaction}: $\Lambda_{ab}$ models the compatibility between fields $a$ and $b$, based on cross-field semantic similarity.
\end{enumerate}

Together, they form:

\[
J_{jk}^{(a,b)} = 
\exp\left( -\frac{|f_j^{(a)} - f_k^{(b)}|}{T} \right)
\cdot 
\exp\left( -|H_j^{(a)} - H_k^{(b)}| \right)
\cdot 
\Lambda_{ab}.
\]

\subsection{Semantic Interaction Matrix $\Lambda$}

The Semantic Interaction Matrix encodes the \textit{meaning-level compatibility} between fields, derived from information geometry:

\[
\Lambda = 
\begin{bmatrix}
1 & \lambda_{12} & \cdots & \lambda_{1F} \\
\lambda_{21} & 1 & \cdots & \lambda_{2F} \\
\vdots & \vdots & \ddots & \vdots \\
\lambda_{F1} & \lambda_{F2} & \cdots & 1
\end{bmatrix}.
\]

Each element $\lambda_{ab}$ is computed from semantic embeddings $\mathbf{e}_a$ and $\mathbf{e}_b$ of field-types:

\[
\lambda_{ab} 
= 
\frac{
\exp\left( \frac{\cos(\mathbf{e}_a, \mathbf{e}_b)}{\tau} \right)
}{
\sum_{c=1}^{F} 
\exp\left( \frac{\cos(\mathbf{e}_a, \mathbf{e}_c)}{\tau} \right)
},
\]

where $\tau$ is a temperature parameter controlling semantic sharpness.  
Thus, $\Lambda$ enables the system to reason using \textbf{meaning}, not only signal-level interactions.

\subsection{Quantum Field Fusion and Structural Evolution}

Fusion and Decomposition now operate not on scalar nodes but on \textit{spinor fields}.  
When entropy concentration is high, SODA-QFT performs:

\[
\Psi_{\text{fused}} = 
\frac{
\sum_{j \in \mathcal{C}} H_j \Psi_j
}{
\sum_{j \in \mathcal{C}} H_j
},
\]

producing higher-order ``field condensates’’ analogous to bound states in quantum many-body systems.

Decomposition similarly splits low-fitness multi-field Nuclei into semantically perturbed offspring:

\[
\Psi_{j,\text{child}} 
= 
\Psi_j + \epsilon \cdot \mathcal{N}(0, \Lambda).
\]

This connects semantic geometry directly to structural dynamics, forming the world’s first \textbf{Semantic Quantum Field Evolution Mechanism}.

\subsection{GPU Fused Kernel Acceleration}

To support multi-field QFT dynamics, the QKM computation is executed using a custom fused CUDA kernel:

\begin{itemize}
    \item Eliminates intermediate tensors for $J_{jk}$ and phase matrices.
    \item Executes cross-field interactions in registers and shared memory.
    \item Reduces global memory traffic by up to $25\times$.
    \item Achieves $12\times$–$24\times$ speedup over PyTorch broadcasting (see Section~\ref{sec:benchmark}).
\end{itemize}

The fused kernel becomes the computational core, enabling SODA-QFT to evolve populations at previously unreachable scales.

\subsection{Interpretation: Meaning as a Physical Interaction}

By introducing $\Lambda$ into the QFT framework, SODA establishes a new physical interpretation:

\begin{quote}
\textit{Meaning is modeled as an interaction field, shaping how distributed computational entities interfere, synchronize, and evolve.}
\end{quote}

This allows SODA-QFT to unify three traditionally separate domains:

\begin{enumerate}
    \item \textbf{Physics}: multi-field quantum interactions.
    \item \textbf{Computation}: GPU-accelerated evolutionary kernels.
    \item \textbf{Semantics}: meaning-driven guidance for structural evolution.
\end{enumerate}

This constitutes a \textbf{Bio-Semantic Quantum Field Theory of Intelligence}, the first of its kind.
